### 2013-09-21 manuscripts/reply/reply.tex
Working on text in reply.
Working on reply.
Also moving `mean_by_dispersion.pdf` plot in the `response-redux`
results.

\newpage
\noindent Recently, this journal has served as a venue for discussion of the
potential pitfalls of approximate Bayesian methods of comparative
phylogeographical model choice.
\citet{Oaks2012} published their findings that the method \msb can often be
biased towards inferring models of temporally clustered divergence times
among taxon pairs.
\citet{Hickerson2013} has published a response to this paper where they
present a model-averaging approach that they conclude circumvents the poor
behavior of the method revealved by \citet{Oaks2012}.
Both of these papers are largely in agreement.
In fact, \citet{Hickerson2013} reiterate a lot of the discussion of
\citet{Oaks2012} regarding the impact of broad uniform priors on Bayesian model
choice.
The main differences in their perspectives are centered around
(1) the means by which broad uniform priors cause the poor behavior of \msb,
and
(2) how to potentially ameliorate this issue.

@@

\section*{Hickerson et al.'s re-analysis of the Philippines data}

\section*{EB}
\citet{Hickerson2013} repeatedly refer the prior distributions used by
\citet{Oaks2012} as ``poorly selected''.
However, this is bit misleading, as \citet{Oaks2012} discuss in detail how they
selected their priors (see their section ``Specifying and simulating the joint
prior'').
They detail how the being limited to uniform prior distributions necessitates
awkward broad priors to avoid excluding the truth, a priori.
What \citet{Hickerson2013} really mean by ``poorly selected'' is that
\citet{Oaks2012} did not use their data to inform their initial prior
distributions (i.e., they did not use an empirical Bayesian approach).
\citet{Oaks2012} did also use empirically informed priors when assessing
both the prior sensitivity of the method and determining the power of \msb
under the ``best'' possible real-world conditions.
\citet{Oaks2012} discuss the potential dangers of taking an empirical Bayesian
model choice (see last paragraph of ``Assessing prior sensitivity of \msb''),
and we expand on this here.

@@

The empirically guided, narrowest prior of \citet{Hickerson2013} ($\divt{} \sim
U(0,0.1)$) likely excludes the true values for at least some of the Philippine
taxa.
The prior matches posterior of \emph{mean} divergence time and is much
narrower than posterior under two of three priors.

@@

\citet{Hickerson2013} translate their results from units of
\globalcoalunit generations to generations assuming a mutation rate of
$1.92\e{-8}$ mutations per site per generation.
If we tranlate there results to a scale more consistant with rates of mutation
of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
by \citet{Hickerson2013} the method can only reliably reject the extreme case
of one divergence event when divergences are random over a period of more than
3 million generations.
Again, considering \citet{Hickerson2013} are assuming prior knowledge that
all population pairs diverged within the past coalescent unit, this
does not seem particularly powerful.

\section*{Lindley's paradox versus sampling error}

\subsection*{The probabilistic argument for sampling error}

@@

\section*{Other clarifications}
\begin{itemize}
    \item saturation plot
    \item different vocabulary (``simultaneous'', etc)
    \item we did not conclude any inference of $\numt{} > 1$ is robust
\end{itemize}

@@

%:TABLE-model-choice-empirical
\begin{table}[htbp]
    \sffamily
    \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{Results of the model-averaging approach of \citet{Hickerson2013}
        applied to the Philippines dataset of \citet{Oaks2012} using three sets
        of prior models. All models used priors on population size of
        $\meanDescendantTheta{} \sim U(0.0001,0.1)$ and $\ancestralTheta{} \sim
        U(0.0001, 0.05)$, and differ only in their prior on divergence time
        (\divt{}) parameters.  Each set of five models differ only in the
        divergence time prior used for the model with the narrowest prior:
        $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_{1A}$ ($\divt{} \sim U(0, 0.01)$),
        or $M_{1B}$ ($\divt{} \sim U(0, 0.001)$). The approximate posterior
        probability of each model ($p(M_i \given \ssSpace)$) is given for each
        of the three analyses.  The posterior estimates are based on 10,000
        samples retained from $1\e6$ prior samples
    from each model.}
    \centering
    \begin{tabular}{ l l l l l }
        \toprule
        & & \multicolumn{3}{c}{$p(M_i \given \ssSpace)$} \\
        \cmidrule(){3-5}
        Model & \divt{} prior & $M_{*}=M_1$ & $M_{*}=M_{1A}$ & $M_{*}=M_{1B}$ \\
        \midrule
        $M_*$ & --        & 0.899 & 0.821 & 0.673 \\
        $M_2$ & $U(0,1)$  & 0.079 & 0.136 & 0.251 \\
        $M_3$ & $U(0,5)$  & 0.013 & 0.026 & 0.044 \\
        $M_4$ & $U(0,10)$ & 0.006 & 0.012 & 0.022 \\
        $M_5$ & $U(0,20)$ & 0.003 & 0.005 & 0.010 \\
        \bottomrule
    \end{tabular}
    \label{tabModelChoiceEmpirical}
\end{table}

@@

\mFigure{../../response-redux/results/hickerson/pymsbayes-results/mean_by_dispersion.pdf}{
    Results of reanalysis.
}{figJointPosterior}

@@

\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
    Sampling error \ldots
}{figSamplingError}

>>>>

\newpage
\noindent Recently, this journal has served as a venue for discussion of the
potential pitfalls of approximate Bayesian methods of comparative
phylogeographical model choice.
The discussion is centered around the approximate Bayesian model choice method,
\msb, which estimates the temporal distribution of divergences among
co-distributed pairs of taxa.
\citet{Oaks2012} published their findings that the method \msb can often be
biased towards inferring models of temporally clustered divergence times
among taxon pairs.
\citet{Hickerson2013} has published a response to this paper where they
present a model-averaging approach that they conclude circumvents the poor
behavior of the method revealved by \citet{Oaks2012}.
Both of these papers are largely in agreement.
In fact, \citet{Hickerson2013} reiterate a lot of the discussion of
\citet{Oaks2012} regarding the impact of broad uniform priors on Bayesian model
choice.
The main differences in their perspectives are centered around
(1) the means by which broad uniform priors cause the poor behavior of \msb,
and
(2) how to potentially ameliorate this issue.

@@

\section*{Hickerson et al.'s re-analysis of the Philippines data}
\citet{Hickerson2013} re-analyze the dataset of \citet{Oaks2012} using a
model-averaging approach, where they place a discrete uniform prior over eight
different prior models (see Table 1 of \citet{Hickerson2013}).
In our initial attempts to explore the behavior of this approach, we found
a fundamental error in the methodology of \citet{Hickerson2013}:
their model mixes different time units.

Each of the eight prior models used in the re-analysis by \citet{Hickerson2013}
has one of two priors on the mean population size of the descendant populations
of each taxon pair:
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ or
$\meanDescendantTheta{} \sim U(0.0005, 0.04)$ or
As described in \citet{Oaks2012}, the divergence time parameters of the model
implemented in \msb are scaled relative to a constant reference population
size, \myTheta{C}.
This reference population size is defined in terms of the upper limit on the
uniform prior on the mean population size of the descendant populations,
\meanDescendantTheta{}, such that for the prior $\meanDescendantTheta{} \sim
U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
$\myTheta{C} = \uniformMax{\meanDescendantTheta{}}/2$.
Thus, the model used by \citet{Hickerson2013} mixes two different units of
time.
In other words, some of their prior and posterior samples are in units of
$0.05/\mutationRate$ generations, whereas other are in units of
$0.02/\mutationRate$ generations.

The fact that their posterior samples are in different units makes the results
of \citet{Hickerson2013} difficult to interpret, and renders their
regression-adjusted results invalid; a fundamental assumption of regression is
that all of the values of the response variable are in the same units.
Thus, the results in section ``Improved Sampling Efficiency by Prior Weighting
Supports Asynchronous and Recent Divergence for the Philippines Vertebrate
Data'' and presented in Figure 2 of \citet{Hickerson2013} should be
disregarded.
The error is easily illustrated by re-plotting their results with the different
time units indicated (Figure~\ref{figJointPosteriorHickerson}.

% In addition to this error, we were not able to reproduce the results of
% \citet{Hickerson2013}.
% We followed their methods, generating $5\e6$ prior samples for each of the
% eight models in their Table 1.
% We retained the 10,000 samples from each model with the smallest Euclidean
% distance to the observed summary statistics, and subsequently retained 10,000
% samples from the remaining 80,000.

\section*{The dangers of empirical Bayesian model choice}
\citet{Hickerson2013} repeatedly refer to the prior distributions used by
\citet{Oaks2012} as ``poorly selected''.
However, this is bit misleading, as \citet{Oaks2012} discuss in detail how they
selected their priors (see their section ``Specifying and simulating the joint
prior'').
They detail how the being limited to uniform prior distributions necessitates
awkward broad priors to avoid excluding the truth, a priori.
What \citet{Hickerson2013} really mean by ``poorly selected'' is that
\citet{Oaks2012} did not use their data to inform their initial prior
distributions (i.e., they did not use an empirical Bayesian approach).
\citet{Oaks2012} did use empirically informed priors when assessing
both the prior sensitivity of the method and determining the power of \msb
under the ``best'' possible real-world conditions.
\citet{Oaks2012} discuss the potential dangers of taking an empirical Bayesian
model choice (see last paragraph of ``Assessing prior sensitivity of \msb''),
and we expand on this here by exploring the theoretical and practical
implications of using an empirical Bayes approach to model choice.

@@

The empirically guided, narrowest prior of \citet{Hickerson2013} ($\divt{} \sim
U(0,0.1)$) likely excludes the true values for at least some of the Philippine
taxa.
They set this prior to match the 95\% highest posterior density (HPD) interval for
the mean divergence time estimated under one of the priors used by
\citet{Oaks2012} (see Table 2 and 3 of \citet{Oaks2012}).
Given this interval estimate is for the \emph{mean} divergence time across all
22 taxa, it seems odd to set this as the prior, as many of the taxon pairs
likely diverged at times older than the upper limit.
Furthermore, the prior is \emph{excluded} from the 95\% HPD interval estimates
of the mean divergence time under the other two prior analyses (under these
priors the 95\% HPD is approximately 0.3--0.6; see Table~6 of
\citet{Oaks2012}).

@@

\citet{Hickerson2013} translate their results from units of
\globalcoalunit generations to generations assuming a mutation rate of
$1.92\e{-8}$ mutations per site per generation.
If we tranlate there results to a scale more consistent with rates of mutation
of nuclear loci ($1\e{-9}$), even under the very optimistic prior settings used
by \citet{Hickerson2013} the method can only reliably reject the extreme case
of one divergence event when divergences are random over a period of more than
3 million generations.
Again, considering \citet{Hickerson2013} are assuming prior knowledge that
all population pairs diverged within the past coalescent unit, this
does not seem particularly powerful.

\section*{Lindley's paradox versus sampling error}
One of the points of disagreement between \citet{Hickerson2013} and
\citet{Oaks2012} is the underlying mechanism causing the bias toward models
with clustered divergences (i.e., models with few divergence events).
\citet{Oaks2012} present simulations based analyses that suggest the broad
uniform prior on divergence time parameters hinder the marginal likelihoods of
models with more divergence time parameters, whereas \citet{Hickerson2013}
argue that the bias is inherent to the inefficient rejection algorithm
implemented in \msb.

\citet{Hickerson2013} present an interesting probabilistic argument in an
attempt to show that insufficient prior sampling is to blame for the bias.
They argue the widest of the three priors used by \citet{Oaks2012} was very
unlikely to produce any prior samples with high numbers of divergence times
that were consistent with the Philippine data.
There are a few issues with their argument.
First, the probabilities they present assume the gene divergence times
estimated by \citet{Oaks2012} are correct.
The sole purpose of the estimates of ultrametric gene trees presented by
\citet{Oaks2012} was to provide a very rough comparison of the gene divergence
times across the 22 taxa.  These analyses assumed an arbitrary strict clock of
$2\e{-8}$ for all taxa, and are, of course, subject to estimation error.
Furthermore, the units of the gene trees are in millions of year, whereas the
divergence time prior of \msb is in generations, thus the logic of
\citet{Hickerson2013} requires the additional assumption that all 22 Philippine
taxa have a generation time of one year.
Thus, the argument of \citet{Hickerson2013} that divergence time estimates of
\citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
dangerous, especially given our findings presented above.

Finally, even \emph{if} we make all of these assumptions and assume the
gene divergence times are without error, the probabilistic argument only
applies to one of the three different priors used by \citet{Oaks2012}.
The narrowest prior on divergence times used by \citet{Oaks2012} closely mirrors
the range of gene divergence time estimates, and applying Hickerson et al.'s
\citeyear{Hickerson2013} probability equations demonstrates that the prior
is densely populated with samples with large number of divergence parameters
that are consistent with the gene divergence estimates
Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
sampling is to blame for the bias, it should be much reduced under the narrow
prior on \divt{}.
However, the magnitude of the bias is very similar across all three priors
\citet{Oaks2012}.
\citet{Hickerson2013} point out a case where the narrow prior performs
slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
However, it is important to note that these results suffered from a bug
in \msb, and there are many cases after \citet{Oaks2012} corrected the 
bug where the narrow prior performs slightly worse (see panels D--J of
Figures 3 and S12).

To disentangle whether Lindely's paradox or insufficient prior sampling is
to blame for the biases revealed by \citet{Oaks2012}, we must look at
the different predictions made by these two phenomena.
One examples, as discussed by \citet{Oaks2012} is that insufficient
prior sampling should create higher variance in posterior estimates,
and thus the analyses should be sensitive to the number of samples
drawn from the prior.
\citet{Oaks2012} do not see such sensitivity when they compare prior sample
sizes of $2\e6$, $5\e6$, and $10\e7$.

To explore this prediction further, we repeat the analysis of the Philippines
dataset under the intermediate prior used by \citet{Oaks2012} ($\divt \sim U(0,
10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
(0.0005, 0.02)$), using a very large prior sample size of $10\e8$.
When we look at the trace of the estimates of the dispersion index of
divergence times (\vmratio{}) as the prior samples accumulate
(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or GLM
regression adjusted estimates.
This strongly suggests that insufficient prior sampling did not play a large
role in the bias found by \citet{Oaks2012}.

Our results presented above that demonstrate the bias of the model-averaging
approach of \citet{Hickerson2013} both toward models with narrower \divt{}
priors (Table~\ref{tabModelChoiceEmpirical} and
Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
\ref{figPowerOmegaProb}) strongly implicate Lindley's paradox as a primary
cause.
In all of our model-averaging analyses, all prior models have the same number
of samples.
Thus, it is difficult to explain the method's propensity towards the model with
the smallest parameter space with insufficient sampling.
While analyses that sample each model proportional to their relative parameter
space could be explored, it seems very likely that the broad uniform priors
are simply inhibiting the marginal likelihoods of spacious models.

As discussed by \citet{Oaks2012}, a straightforward prediction of Lindley's
paradox is that it should disappear as the model generating the data converges
to the prior.
\citet{Oaks2012} test this prediction by performing 100,000 simulations to
assess the model-choice behavior of \msb when the prior model is correct.
They find that the bias actually switches direction (at least for the
regression-adjusted estimates) and the method tends to underestimate the
probability of the model with one divergence event (see Figure 4 of
\citet{Oaks2012}).
The same prediction is not as straightforward for the insufficient-sampling
hypothesis.
Even when the prior is correct, due to the discrete uniform prior on
\numt{} implemented in \msb, models with larger numbers of divergence
events (and thus greater parameter space) will still be less densely
sampled than those with fewer divergence events \citep{Oaks2012}.

@@

\section*{Other clarifications}

\subsection*{saturation plot}
\citet{Hickerson2013} claim that our priors ``cause much of the explored
parameter space to be beyond the threshold of saturation in most mtDNA genes.''
To explore this possibility, we simply plotted the summary statistics used
over divergence times to look for the point when the statistics begin to
saturate (Figure~\ref{figSaturationPlot}).
Clearly, the priors with upper limits on \divt{} of 5 and 10 did not
suffer from saturation.
Even as far out as divergence times as 20 coalescent units, while 
there is some saturation, there is still clearly signal in the
summary statistics used by \msb.

\begin{itemize}
    \item different vocabulary (``simultaneous'', etc)
    \item we did not conclude any inference of $\numt{} > 1$ is robust
\end{itemize}

@@

%:TABLE-re-analysis
% \begin{table}[htbp]
%     \sffamily
%     \footnotesize
%     \addtolength{\tabcolsep}{-0.08cm}
%     \rowcolors{2}{}{myGray}
%     %\captionsetup{font=footnotesize}
%     \caption{Results of the model-averaging approach of \citet{Hickerson2013}
%         applied to the Philippines dataset of \citet{Oaks2012} using eight
%         prior models (see \citet{Hickerson2013} Table 1).  The results of (1)
%         \citet{Hickerson2013}, (2) our full re-analysis, and of (3) repeating
%         the final rejection step of \citet{Hickerson2013} using the prior means and
%         standard deviations to normalize the summary statistics, are all compared.}
%     \centering
%     \begin{tabular}{ l l l l }
%         \toprule
%          & Hickerson et al. (2013) & Re-analysis & blah \\
%         $\hat{\numt{}}$ & 2 & 1 & 1 \\
%         $p(\numt{} = 1 | \ssSpace)$ & 0.062 & 0.206 & 0.146 \\
%         $BF_{\numt{} = 1, \numt{} \neq 1}$ 1.39 & 5.45 & 3.59 \\
%         $p(\vmratio{} < 0.01 | \ssSpace)$ & 0.298 & 0.416 & 0.361 \\
%         $BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01}$ 6.65 & 11.16 & 8.85 \\
%         $\hat{M}$ & 1 & 2 & 2 \\
%         $p(M_1 | \ssSpace)$ & 0.513 & 0.358 & 0.376 \\
%         $p(M_2 | \ssSpace)$ & 0.286 & 0.369 & 0.389 \\
%         $p(M_3 | \ssSpace)$ & 0.036 & 0.086 & 0.089 \\
%         $p(M_4 | \ssSpace)$ & 0.002 & 0.012 & 0.011 \\
%         $p(M_5 | \ssSpace)$ & 0.142 & 0.129 & 0.107 \\
%         $p(M_6 | \ssSpace)$ & 0.020 & 0.043 & 0.024 \\
%         $p(M_7 | \ssSpace)$ & 0.001 & 0.003 & 0.004 \\
%         $p(M_8 | \ssSpace)$ & 0.000 & 0.000 & 0.001 \\
%         \bottomrule
%     \end{tabular}
%     \label{tabModelChoiceEmpirical}
% \end{table}

%:TABLE-model-choice-empirical
\begin{table}[htbp]
    \sffamily
    % \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{Results of the model-averaging approach of \citet{Hickerson2013}
        applied to the Philippines dataset of \citet{Oaks2012} using three sets
        of prior models. All models used priors on population size of
        $\meanDescendantTheta{} \sim U(0.0001,0.1)$ and $\ancestralTheta{} \sim
        U(0.0001, 0.05)$, and differ only in their prior on divergence time
        (\divt{}) parameters.  Each set of five models differ only in the
        divergence time prior used for the model with the narrowest prior:
        $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_{1A}$ ($\divt{} \sim U(0, 0.01)$),
        or $M_{1B}$ ($\divt{} \sim U(0, 0.001)$). The approximate posterior
        probability of each model ($p(M_i \given \ssSpace)$) is given for each
        of the three analyses.  The posterior estimates are based on 10,000
        samples retained from $1\e6$ prior samples
    from each model.}
    \centering
    \begin{tabular}{ l l l l l }
        \toprule
        & & \multicolumn{3}{c}{$p(M_i \given \ssSpace)$} \\
        \cmidrule(){3-5}
        Model & \divt{} prior & $M_{*}=M_1$ & $M_{*}=M_{1A}$ & $M_{*}=M_{1B}$ \\
        \midrule
        $M_*$ & --        & 0.899 & 0.821 & 0.673 \\
        $M_2$ & $U(0,1)$  & 0.079 & 0.136 & 0.251 \\
        $M_3$ & $U(0,5)$  & 0.013 & 0.026 & 0.044 \\
        $M_4$ & $U(0,10)$ & 0.006 & 0.012 & 0.022 \\
        $M_5$ & $U(0,20)$ & 0.003 & 0.005 & 0.010 \\
        \bottomrule
    \end{tabular}
    \label{tabModelChoiceEmpirical}
\end{table}

@@

% \mFigure{../../response-redux/results/hickerson/pymsbayes-results/pymsbayes-output/d1/m12345678-combined/mean_by_dispersion.pdf}{
%     Results of reanalysis.
% }{figJointPosterior}

@@

\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
}{figSamplingError}
